Online Verification of Commutativity

### 2020-08-02 acmPaper/paper.tex
Evaluation draft

\section{Evaluation}
We compare performance of the following path checking algorithms: (1) naive baseline (2) less naive, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
The two aspects we look at are time for response and size of response set (smaller sets -- tighter output results -- would mean less calls to the oracle).
We use randomly generated graphs for benchmarking, and vary their size.
All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

The naive baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude.
All implementations are sensitive to density, performing better when density is very low or very high.

%TODO add figures

\begin{tabular}{|c|c|}
    \hline
    Algorithm & average seconds taken for ten nodes, density close to 0.5 \\
    \hline
    Naive baseline & 1,500 \\
    Batch algorithm & 1.4 \\
    algorithm ~\ref{algo_online_polynomial} & 1.2 \\
    algorithm ~\ref{algo_online_minimal} & 0.0006 \\
    \hline
\end{tabular}

>>>>

\section{Evaluation}
We compare performance of the following path checking algorithms: (1) naive baseline (2) less naive, two-flip baseline (3) batch baseline (4) algorithm ~\ref{algo_online_polynomial}, and (5) algorithm ~\ref{algo_online_minimal}.
The two aspects we look at are time for response and size of response set (smaller sets -- tighter output results -- would mean less calls to the oracle).
We use randomly generated graphs for benchmarking, and vary their size.
Given a graph and a new edge, we time how long it takes for an algorithm to return 
All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

The naive baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
The batch algorithm brings timing down to seconds. Algorithm ~\ref{algo_online_polynomial} performs only slightly better.
Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
All implementations are sensitive to density, performing better when density is very low or very high.

\subsection{Comparison of algorithm time cost}
The average time taken by each algorithm over the course of 10 runs over randomly generated graphs with 9 nodes and 32 edges is listed in table ~\ref{times}.

\begin{figure}
\begin{tabular}{|c|c|}
    \hline
    Algorithm & average seconds of computation \\
    \hline
    Naive baseline & 0.77 \\
    Two Flip tolerant & 0.075 \\
    Batch algorithm & 7.55 \\
    ~\ref{algo_online_polynomial} & 0.0038 \\
    ~\ref{algo_online_minimal} & 0.00086 \\
    \hline
\end{tabular}
\label{times} %TODO debug and fix label not working.
\caption{Computation time for 9 node graph of density 0.4, averaged over ten runs}
\end{figure}

\subsection{Scaling of time with input size}

Algorithms scale as expected with time, as seen in figure ~\ref{fig:timeVsSize}. 
Both algorithm ~\ref{algo_online_minimal} and ~\ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance, as does the batch baseline.
\xxx[AK]{Do I need to justify this by doing some kind of fit?}
The naive baseline as well as the two flip tolerant baseline display growth that quickly makes computation untenable.
The batch algorithm also grows quickly.

We define density to be the ratio of the number of edges in the graph to the total possible number of edges ($|V|^2$, where $|V|$ is the number of nodes).
Run time relates to the density of edges in the input graph.
The degree of the effect differs with the algorithms, as figure ~\ref{fig:timeVsSize} shows.
Generally, denser graphs entail longer computation time.
% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.

\begin{figure}
    % TODO format nicely.
    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{timeVsSize_10_OptimalSet.png}
      \caption{~\ref{algo_online_minimal}}
      \label{fig:sfigOptimal}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{timeVsSize_10_Polynomial.png}
      \caption{~\ref{algo_online_polynomial}}
      \label{fig:sfigPolynomial}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{timeVsSize_10_NaiveChecker.png}
      \caption{Naive baseline}
      \label{fig:sfigNaive}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{timeVsSize_10_TwoFlipPathChecker.png}
      \caption{Two flip tolerant baseline}
      \label{fig:sfigTwoFlip}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{timeVsSize_10_BatchChecker.png}
      \caption{Batch algorithm baseline}
      \label{fig:sfigBatch}
    \end{subfigure}

    \label{fig:timeVsSize}
    \caption{Plots of scaling of time with input size and density}
\end{figure}

\subsection{Variance}
The periodic spikes in figure ~\ref{fig:sfigOptimal} are striking. 
We plot the spread of results to understand what is happening.
We find ~\ref{algo_online_minimal} has outliers about two standard deviation above the mean responsible for the spikes in the average.
The outliers themselves follow a polynomial curve.
The plots in figure ~\ref{fig:variance} depict the situation.
Grey points are the results of evaluation on individual points, and error bars show standard deviation. 
The black curve traces the mean.
No such effect is observed in ~\ref{algo_online_polynomial}.
\begin{figure}
    % TODO format nicely.
    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{variance_10_OptimalSet.png}
      \caption{~\ref{algo_online_minimal}}
      \label{fig:sfigOptimal}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{variance_10_Polynomial.png}
      \caption{~\ref{algo_online_polynomial}}
      \label{fig:sfigPolynomial}
    \end{subfigure}

    \begin{subfigure}{.4\textwidth}
      \includegraphics[width=.8\linewidth]{variance_10_TwoFlip.png}
      \caption{Two flip tolerant baseline}
      \label{fig:sfigTwoFlip}
    \end{subfigure}

    \label{fig:variance}
    \caption{Plots of spread of results}
\end{figure}

\subsection{Size of Output}
Output size is a metric of interest, should the equality checking oracle be expensive.
Table ~\ref{sizes}, summarizes the number of output pairs that the algorithms returned on average over 10 runs, for graphs with 9 nodes and 32 edges.
\begin{figure}
\begin{tabular}{|c|c|}
    \hline
    Algorithm & average number of output pairs \\
    \hline
    Naive baseline & 39754.9 \\
    Two Flip tolerant & 748.9 \\
    Batch algorithm & 23 \\
    algorithm ~\ref{algo_online_polynomial} & 78.3 \\
    algorithm ~\ref{algo_online_minimal} & 1 \\
    \hline
\end{tabular}
\label{sizes} % Now working. TODO fix label.
\caption{Output size for 9 node graph of density 0.4, averaged over ten runs}
\end{figure}
